Discrepancy, Coresets, and Sketches in Machine Learning

### 2019-01-09 komlos.tex
more related work

>>>>

\section{Related Work}
\begin{itemize}
\item coresets - motivation, some papers, streaming algorithms for coresets

\item Radamacher complexity \cite{Bartlett:2003:RGC:944919.944944} is a known method to estimate the generalization ability of an empirical risk minimizer (ERM) of a family of function over a finite set of i.i.d sampled points. A carefully selected set of points can potentially obtain a better generalization bound, and that is the usefulness of coresets in the context of machine learning. We are not aware of any analog to Radamacher complexity that aims to measure the generalization ability of a coreset of a fixed finite size. There are papers such as \cite{tolochinsky2018coresets} and references within that tie the coreset to the VC dimension of the family function, and the average sensitivity of the dataset. However, these relationships come up more as tools for constructing a coreset rather than a single complexity measure aimed to characterize some generalization ability.
%
% There is literature about the complexity of active learning for the supervised setting but that is out of scope as we deal with the setting where all labels are available, or the problem is unsupervised. For the complexity of coresets, the total sensitivity of the dataset and the VC dimension of the function class is often used as tools for construction \zk{cite something?}. However, to the best of our knowledge there is no clear characterization of the size of coresets based on some complexity measure of the dataset and function class.

\item Radamacher complexity \cite{Bartlett:2003:RGC:944919.944944}, There is literature about the complexity of active learning for the supervised setting but that is out of scope as we deal with the setting where all labels are available, or the problem is unsupervised. For the complexity of coresets, the total sensitivity of the dataset and the VC dimension of the function class is often used as tools for construction \zk{cite something?}. However, to the best of our knowledge there is no clear characterization of the size of coresets based on some complexity measure of the dataset and function class.

\item Kernel density estimation is a popular tool in data analysis aimed to estimate a continuous distribution with a finite set of points. Among other applications, this tool is used for outlier detection \cite{schubert2014generalized}, regression \cite{fan2018local}, and clustering \cite{rinaldo2010generalized}. A thorough survey could be found in \cite{silverman2018density}. Given a set of $n$ data points $x_1,\ldots,x_n$ and a query $q$, the density estimate of $q$ is the average $\frac{1}{n} \sum_i K(x_i,q)$ of the kernel evaluations with the data points. The task of finding a coreset for kernel density estimates consists of obtaining a set of points $z_1,\ldots,z_m$ with weights $w_1,\ldots,w_m$ for some $m \ll n$ such that for any query $q$ it holds that 
$\frac{1}{n} \sum_i K(x_i,q) = \frac{1}{m} \sum_i w_i K(z_i,q) \pm \eps$. For this task, the state-of-the-art is given by \cite{DBLP:journals/corr/abs-1802-01751}, achieving $m=\sqrt{d\log(1/\eps)}/\eps$ where $d$ is the dimension of the original data points. Their result holds for Lipchitz psd kernels. The result is constructive though it is polynomial rather than (quasi-)linear in the data size. The authors give an almost matching lower bound of $\sqrt{d}/\eps$ and pose an open question for closing the gap between the bounds.

\item Logistic Regression is an extremely popular technique for classification task. In it, we are given a set of $n$ pairs $x_i,y_i$ with $x_i \in \R^d$ and $y_i \in \{-1,1\}$. The objective is to find a regressor $q \in \R^d$ minimizing the logistic loss defined as $\E_i \log(1+\exp(x_i^\top q y_i))$. A coreset for logistic regression is a set of $m \ll n$ pairs $z_i, v_i$ in $\R^d, \{-1,1\}$ correspondingly along with weights $w_i$, with the property that for any regressor $q$, either $\E_i \log(1+\exp(x_i^\top q y_i)) = \E_i \log(1+\exp(z_i^\top q v_i)) \pm \eps$ or $\E_i \log(1+\exp(x_i^\top q y_i)) = \E_i \log(1+\exp(z_i^\top q v_i)) (1 \pm \eps)$. A recent paper \cite{DBLP:journals/corr/abs-1805-08571} provides a coreset with a multiplicative guarantee that is based on an average sensitivity property of the dataset. They provide a lower bound for the size of a multiplicative coreset showing that in general it is not possible to achieve $m \ll n$. The coreset they build is of cardinality $m \approx \mu\sqrt{nd^3}/\eps^2$ where $\mu \geq 1$ is complexity measure of the dataset.  \cite{tolochinsky2018coresets} gives a generic multiplicative coreset construction for any monotonic function with $\ell_2^2$ regularization. The dependence they get is, ignoring logarithmic factors, is $d/\eps^2$.

%\item coreset of logistic regression  \cite{DBLP:journals/corr/abs-1805-08571}, \cite{DBLP:conf/nips/HugginsCB16}. Give multiplicative approximation. This is impossible in general so they depend on a different restriction on the regressor or dataset. Based on sensitivity scores. \cite{DBLP:conf/nips/HugginsCB16} use a heuristic of using $k$-means to approximate sensitivity scores and prove that given good sensitivity scores they get a good coreset. \cite{tolochinsky2018coresets} gives a generic coreset construction for any monotonic function as long as it comes with an $\ell_2^2$ regularization. The dependence they get is, ignoring polynomial factors, $d/\eps^2$.  \cite{DBLP:journals/corr/abs-1805-08571} have a different method, relying on a measure of complexity of the dataset and obtain roughly $\sqrt{n}d^{3/2}/\eps^2$ sized dataset. 
\end{itemize}
